---
title:     Identification of Periods of Clear Sky Irradiance in Time Series of GHI Measurements Matthew J. Reno and Clifford W. Hansen.
author:    Natsis Athanasios
documentclass: article
classoption:   a4paper,oneside
fontsize:      10pt
geometry:      "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"

link-citations:  yes
colorlinks:      yes

header-includes:
  - \usepackage{caption}
  - \usepackage{placeins}
  - \captionsetup{font=small}
  - \usepackage{multicol}
  - \setlength{\columnsep}{1cm}

output:
  bookdown::pdf_document2:
    number_sections:  yes
    fig_caption:      no
    keep_tex:         no
    toc_depth:        4
    latex_engine:     xelatex
    toc:              yes
    fig_width:        7
    fig_height:       4.5
---

```{r setup, echo=F, include=T, eval=TRUE}
MS <- data.frame(readRDS("PARAMS/Clear_sky_id_Reno-Hansen_apply_v14_2.Rds"))
# export all variables
for (an in names(MS)) {
    assign(an, MS[[an]])
}
```



<!-- `````{r echo=F, include=F, eval=F} -->
<!-- knitr::opts_chunk$set(comment    = ""      ) -->
<!-- # knitr::opts_chunk$set(dev        = "pdf"   ) -->
<!-- knitr::opts_chunk$set(dev        = "png"    ) -->
<!-- knitr::opts_chunk$set(out.width  = "100%"   ) -->
<!-- knitr::opts_chunk$set(fig.align  = "center" ) -->
<!-- # knitr::opts_chunk$set(fig.pos    = '!h'    ) -->


<!-- #  Use the optimized alpha for the selected models and identify Clear-Sky -->
<!-- #  minutes. -->

<!-- ## load previous state have to override it for alpha to be used -->
<!-- load("~/Aerosols/DATA/model_opt/Combinations_results_2022-06-14_153313.Rds") -->


<!-- ## Define the model to use for CS detection be inspecting the results of the 'a' optimization script. -->
<!-- model_selection <- "HAU" -->
<!-- monthly_alphas  <- gather_results[gather_results$CS_models == model_selection,] -->

<!-- ## daily plots file name -->
<!-- plotsbase <- paste0("~/CS_id/REPORTS/DAILY/", -->
<!--                     sub("\\.R$", "", basename(Script.Name)), "_") -->
<!-- ````` -->





\newpage

# Detection of clear periods in GHI measurements for SDR tredns.


<!-- ## Only filters for GHI are used for SDR trends -->

<!-- ## Load all data from `DATA/Broad_Band/QCRad_LongShi/QCRad_LongShi_v8_apply_CM21_CHP1_[0-9]{4}.Rds` -->

<!-- ##  Exclude data where `wattGLB < wattHOR` -->

<!-- There are instances where global irradiance is less than direct. -->

<!-- This happens, near sunset and sunrise due to obstacles, -->
<!-- or due to different sunrise/sunset time due to small spatial differences. -->
<!-- We will exclude all this data both for global and direct. -->


<!-- ##  Exclude other bad data identified in raw data -->






### Threshold Values for Criteria

For mean and max Most evaluations of clear sky models find that the average
bias error of the model is less than 10%, often around 7% [1, 83]. Therefore,
a fixed threshold of $\pm 75 W / m^2$ within the mean and max of the clear sky model
was chosen.



### Clear Sky detection Algorithm values


<!-- `````{r include=F, eval=F, echo=F} -->
<!-- # panderOptions('table.emphasize.rownames', F) -->
<!-- panderOptions('table.alignment.default',  "right") -->
<!-- panderOptions('table.alignment.rownames', "right") -->
<!-- panderOptions('table.split.cells',        c(50,10)) -->
<!-- ````` -->




### Reference clear sky irradiance model and air mass model (AM)

We select the clear sky model, the air mass calculation
and the period of $`r nt`$ minutes to use.
The characterization is done to the center value of the window.


`````{r include=F, eval=F, echo=F}
AM       <- AM_simple
CS_model <- get(model_selection)
`````


`````{r include=F, eval=T, echo=F}
MS <- data.frame()
`````

Air mass model

$$ AM = \frac{1}{\cos(z)} $$

### Clear Sky irradiance Haurwitz model (HAU) (1945)  

\begin{equation}
\text{GHI}_\text{Clear Sky} = `r signif(alpha,3)` \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right)  (\#eq:ahau)
\end{equation}


## Clear sky Id procedure



<!-- MeanVIP -->
### 1. Mean value of irradiance during the time period 

Comparing value in a moving window of $`r nt`$ minutes.
The mean of the measured value $\overline{G}_i$ inside an envelope based on the reference model $\text{GHI}_\text{Clear Sky}$.

\begin{equation}
`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{GHI}}_{i\text{Clear Sky}} - `r CS_ref_rm_VIP_LowOff`
< \overline{G}_i <
`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{GHI}}_{i\text{Clear Sky}} + `r CS_ref_rm_VIP_UppOff`
(\#eq:MeanVIP)
\end{equation}


<!-- MaxVIP -->
### 2. Max value of irradiance during the time period

The running max measured value $M_{Gi} = max[\text{GHI}_{i}]$ is compared to a similar constructed value from the reference $M_{CSi} = max[\text{GHI}_{i\text{Clear Sky}}]$.

\begin{equation}
`r MaxVIP_fct` \times M_{CSi} - `r MaxVIP_off_low`
< M_{Gi} <
`r MaxVIP_fct` \times M_{CSi} + `r MaxVIP_off_upp`
(\#eq:MaxVIP)
\end{equation}


<!-- VIL -->
###  3. Variability in irradiance by line length

 The length L of the sequence of line segments connecting the points of the GHI time
series is calculated by 

\begin{equation}
L = \sum_{i=1}^{n-1}\sqrt{\left ( \text{GHI}_{i+1} - \text{GHI}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
(\#eq:VILeq)
\end{equation}

for the measured values $L$ and similar $L_{CS}$ using the reference model $\text{GHI}_{i\text{Clear Sky}}$. The criterion is applied

\begin{equation}
`r MinVIL_fct` \times L_{CSi} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` \times L_{CSi} + `r offVIL_upl`
(\#eq:VILcr)
\end{equation}


<!-- VCT -->
### 4. Variance of Changes in the Time series

Standard deviation of rate of change in irradiance. We calculate the standard deviation  of the slope
(s) between sequential points in the time series, normalized by the average GHI during the time interval:

\begin{gather}
s_i = \frac{\text{GHI}_{i+1} - \text{GHI}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} (\#eq:VCT1) \\
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i (\#eq:VCT2) \\
\sigma = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G}} (\#eq:VCT3)
\end{gather}

\begin{equation}
\sigma_i < `r offVCT`
(\#eq:VCTcr)
\end{equation}



<!-- VSM -->
### 5. Variability in the Shape of the irradiance Measurements 

Maximum difference between changes in GHI and clear sky time series. The final quantity considered
is the maximum difference $X$ between the change in measured irradiance and the change in clear sky
irradiance over each measurement interval:

\begin{gather}
x_i = \text{GHI}_{i+1} - \text{GHI}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} (\#eq:VSM1) \\
x_{CS,i} = \text{GHI}_{CS,i+1} - \text{GHI}_{CS,i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} (\#eq:VSM2) \\
X_i = \max{\left \{ \left | x_i - x_{CS,i} \right | \right \}} (\#eq:VSM3) 
\end{gather}

\begin{equation}
X_i < `r offVSM`
(\#eq:VSMcr)
\end{equation}


<!-- IGNORED FOR GHI TRENDS -->
<!-- LDI --> 
<!-- ### 6. Low Direct Irradiance limit (LDI) -->

<!-- Ignore Direct irradiance when it is bellow $5 Watt/m^2$. -->
<!-- This limit is biased due to the slight location difference of the two Instruments. -->
<!-- The difference in shadow especially the afternoon "pole shadow" of some periods of the year. -->
<!-- May limit the implementation by time of day or/and period of the year -->

<!-- Also this can not characterize all global measurements due to gaps on direct measurements. -->


<!-- IGNORED FOR GHI TRENDS -->
<!-- LGI  -->
<!-- ### 7. Low Global Irradiance limit (LGI) -->

<!-- Global irradiance below `r VGIlim` $Watt/m^2$ level can not be identified -->


<!-- IGNORED FOR GHI TRENDS -->
<!-- FCS -->
<!-- ### 8. Too Few CS point for the day (FCS) -->

<!-- If in a day there less than 11 clear sky points will exclude them from optimizing -->


<!-- FDP -->
### 9. Too Few data points for the day

Days with less than $`r nt * 3`$ data points (minutes) will be ignored.




<!-- IGNORED FOR GHI TRENDS -->
<!-- DsT -->
<!-- ### 11. Too low Direct signal (DsT) -->

<!-- This is a simple hard limit on the lower possible Direct radiation with clear sky. -->

<!-- The threshold is 25% lower than a reference value. -->

<!-- $$ I_d = I_0 * 0.7^{{AM}^{0.678}} * cos({ZSA}) $$ -->

<!-- Where ${AM}$ is the selected air-mass model. -->

<!-- "Clear sky direct normal irradiance estimation based on adjustable inputs and error correction_Zhu2019.pdf" -->






<!-- ### Baseline value for direct irradiance -->

<!-- See: Clear sky direct normal irradiance estimation based on adjustable inputs and error correction_Zhu2019.pdf -->



| CS Flag | Test |
|:-------:|:--------------------------------------------------------------|
|   NA    | Undefined, untested                                           |
|    0    | Passed as clear sky                                           |
|    1    | Mean value of irradiance during the time period (MeanVIP)     |
|    2    | Max Value of Irradiance during the time Period (MaxVIP)       |
|    3    | Variability in irradiance by the length (VIL)                 |
|    4    | Variance of Changes in the Time series (VCT)                  |
|    5    | Variability in the Shape of the irradiance Measurements (VSM) |
|    6    | Low Direct Irradiance limit (LDI)                             |
|    7    | Low Global Irradiance limit (LGI)                             |
|    8    | Too Few CS point for the day (FCS)                            |
|    9    | Too Few data points for the day                               |
|   10    | Missing Data                                                  |
|   11    | Direct irradiance simple threshold                            |










### Cost function for optimization of alpha value.

$$ f(a) = \dfrac{ \sum_{i=1}^{n} ( a \cdot {GHI}_i - {CSI}_i )^2 }
                { n } , \qquad a > 0 $$





**END**

